---
title: "Day 8 Practice"
output: html_document
---

# Section 1

1. Write an if...else statement to test whether a string matches "right", and if so, print "you win!", but if the string matches "wrong", print "you lose!", and if it's neither, print "the result is inconclusive".

```{r}

x <- "Poop"
if(x == "Right") {
  print("You Win!")
}  else if(x == "Wrong") {
  print("You Lose!")
}  else {
  print("The result is inconclusive")
}

```


2. Make a copy of the iris data. Use `ifelse()` to create a new column, where:
  - if the value of Sepal.Length is less than 5, the new column should have a value of "low"
  - if the value of Sepal.Length is greater than 6, it should have a value of "high"
  - and if it's between 5 and 6 (inclusive), it should have a value of "mid"
  - make a plot in ggplot2 to confirm that you did this right
  
```{r}
library(ggplot2)
my.iris <- iris

x <- my.iris$Sepal.Length
my.iris$Sepal.Lenght2 <- ifelse(x < 5, "Low",
       ifelse(x > 6, "High", "Mid"))
head(my.iris)

factor(my.iris$Sepal.Lenght2)
ggplot(my.iris, aes(Sepal.Lenght2)) + geom_bar(aes(fill = Species))

ggplot(my.iris, aes(Sepal.Length)) + geom_histogram(aes(fill = Sepal.Lenght2))

my.iris$Sepal.Lenght2 <- factor(my.iris$Sepal.Lenght2, labels = c("Low", "Mid", "High"))

ggplot(my.iris) + geom_bar(aes(Sepal.Lenght2, Sepal.Length, fill = Species), position = "dodge", stat = "summary", fun.y = "mean")
```


# Section 2

3. Practice writing a very simple loop, that simply prints out each level of Species from the iris data.  In other words, the result of running the loop should look something like:

"Level number 1 of Species is setosa"
"Level number 2 of Species is versicolor"
"Level number 3 of Species is virginica"

```{r}

#what we want:
#print(paste("Level number", SPECIES NUM, "of Species is", "SPECIES"))
#how do you loop species and species number

levels(iris$Species)

for(levelnum in 1:length(levels(my.iris$Species))) { #dim is dimension
  print(levelnum)
}
for(levelnum in 1:length(levels(my.iris$Species))) { 
  print(paste("Level number", levelnum, "of Species is")) #Level number this.dim
}

for(levelnum in 1:length(levels(my.iris$Species))) { 
  this.level <- levels(iris$Species)[levelnum]
  print(paste("Level number", levelnum, "of Species is", this.level)) #Level number this.dim
}

for(this.level in levels(my.iris$Species)) { 
  levelnum <- which(levels(iris$Species) %in% this.level)
  print(paste("Level number", levelnum, "of Species is", this.level)) #Level number this.dim
}
```

4. Now using the loop from #3 as a starting point, for each pass through the loop, save a PDF of a scatterplot of Sepal.Length by Sepal.Width, but the plot should be different each time through the loop in the following ways:
  - each plot should only plot the subset of data from one species
  - each plot should display the plotted species as the plot title (either use `labs()` or `ggtitle()`)
  - each plot should be saved as a separate file, with the species in the file name
  - how could you change this code to write all three plots out to the same file?

```{r}
# ggtitle() example
ggplot(iris, aes(Sepal.Width)) + geom_histogram() + ggtitle("All species")

#do something once
pdf("all species plot.pdf")
ggplot(iris, aes(Sepal.Width)) + geom_histogram() + ggtitle("All Species")
dev.off()

#construct the loop

for(levelnum in 1:length(levels(my.iris$Species))) { 
  this.level <- levels(iris$Species)[levelnum]
  #print(paste("Level number", levelnum, "of Species is", this.level))
  this.data <- iris[iris$Species %in% this.level, ]
  pdf(paste(this.level, "plot.pdf"))
  print(ggplot(this.data, aes(Sepal.Width)) + geom_histogram() + ggtitle(this.level))
  dev.off()
}

#multiple page pdf

pdf("species plots.pdf")

for(levelnum in 1:length(levels(my.iris$Species))) { 
  this.level <- levels(iris$Species)[levelnum]
  #print(paste("Level number", levelnum, "of Species is", this.level))
  this.data <- iris[iris$Species %in% this.level, ]
  print(ggplot(this.data, aes(Sepal.Width)) + geom_histogram() + ggtitle(this.level))
  dev.off()
}
```

# Section 2

4. Similar to what we did with the correlation power simulations, write code (including a loop) that creates a power simulation for comparing the means of two normally-distributed samples.
  - compare the sample with both `t.test()` and `wilcox.test()`
  - put the p-values from both tests in each simulation result
  - compare the power of the t-test vs. the power of the Wilcoxon test
  
```{r}
# set up the parameters of the simulation
n.sim <- 1000
# what other parameters are needed?

# set up the "empty" data frame for the results
sim.results <- data.frame(# WHAT SHOULD YOUR COLUMNS BE? 
                          )

# start the loop
# each iteration of the loop should:
# 1. generate data
# 2. perform the analysis
# 3. store the results (here, p-values) in the results data frame


# examine the results
# create a column that indicates which of the p-values reached significance


# power = the proportion of "significant" results


```

